Rectangulotopes

### 2024-04-03 draft/rectangulotopes.tex
sect 3.4

Weak rectangulotopes are deformed permutahedra, hence are defined by a \defn{submodular} set function $f:2^{[n]}\mapsto \R$. The description of the weak rectangulotope $\WRP(n)$ as a Minkowski sum in Lemma~\ref{lem:weakMinkowski} allows us to easily compute the submodular function defining it.

\begin{lemma}
  The weak rectangulotope $\WRP(n)$ is realized as
  \[
  \{x\in\R^n : \sum_{i\in S} x_i \leq f(S) \text { and } \sum_{i\in [n]} x_i = f([n])\},
  \]
  where $f:2^{[n]}\mapsto \R$ is the following submodular function:
  \[
  f(S) := |\{ I\text{ interval of } [n] : I\cap S\not=\emptyset \text{ and } I\cap ([n]\setminus S) \not= \emptyset \}|.
  \]
  %that maps $S$ to the number of intervals $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$. 
\end{lemma}

@@

\begin{proof}
  From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}. For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have
  \[
  \SP(A) = \{ x\in\R^n : \sum_{i\in S} x_i \leq \loday{f}_I(S) \text { and } \sum_{i\in [n]} x_i = \loday{f}_I([n]) \},
  \]
  where
  \[
  \loday{f}_I(S) := \llbracket S\cap I\not=\emptyset \rrbracket .
  \]
  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
  Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
  \[
  \antiloday{f}_I(S) := - \llbracket I\subseteq S \rrbracket .
  \]
  Their sum
  \[
  f(S) := \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
  \]
  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
\end{proof}

>>>>

Recall that every deformed permutahedra can be defined as
  \[
  \{x\in\R^n : \sum_{i\in S} x_i \leq f(S)\ \forall S\subseteq [n], S\not=\emptyset, \text { and } \sum_{i\in [n]} x_i = f([n])\},
  \]
where $f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$.
The description of the weak rectangulotope $\WRP(n)$ as a Minkowski sum in Lemma~\ref{lem:weakMinkowski} allows us to easily compute the submodular function defining it.

\begin{lemma}
  The weak rectangulotope $\WRP(n)$ is realized by the following submodular function:
  \[
  f(S) := |\{ I\text{ interval of } [n] : I\cap S\not=\emptyset \text{ and } I\cap ([n]\setminus S) \not= \emptyset \}|.
  \]
  %that maps $S$ to the number of intervals $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$. 
\end{lemma}

@@

\begin{proof}
  From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}. For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
  \[
  \loday{f}_I(S) := \llbracket S\cap I\not=\emptyset \rrbracket .
  \]
  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
  Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
  \[
  \antiloday{f}_I(S) := - \llbracket I\subseteq S \rrbracket .
  \]
  Their sum
  \[
  f(S) := \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
  \]
  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
\end{proof}
